Journey: A Study of Scalabilty of Memory Management in Linux

### 2017-05-05 paper.tex
addressed most of Karan's concerns

For demand paging to work efficiently, the correct \texttt{vm\_area\_struct}
needs to be efficient to find. Linux caches the most recently used region and
organizes all regions into a red-black tree for fast lookup.  Moreover,
adjacent regions with similar permissions and properties are merged in an
effort to keep bookkeeping costs low.

In our first set of experiments, we measure the memory overhead and latency of
the Linux Memory Management system's use of \texttt{vm\_area\_struct}, which
represents a region of a process's virtual address space. All the physical
pages in the system are managed using the \texttt{struct page} descriptors in
the kernel.  We do not measure the overhead of \texttt{struct page} because it
is known that there is one struct for each physical page. We do not measure the
overhead of page tables because this can be calculated from the size of the
virtual memory space used by the process.

@@

The specifications of the machine we used to run our experiments are listed in
Figure \ref{fig:specs}. We disabled Intel SpeedStep and TurboBoost and ensure
that the processor frequency stays constant. All experiments are pinned to the
same core on our machine.

We run several different microbenchmarks, each of which has a different memory
allocation pattern. These benchmarks are designed to stress the memory
management system in different ways, rather than simulate real workloads. 
Each run is executed in isolation, without any additional processes
running beyond services that start when the system is booted and a single ssh
session. We are careful to avoid opening extra ssh sessions, running screen or
tmux, or executing extra commands while a benchmark is running since these may
impact memory usage and, thus, our results.

@@

\subsubsection{Memory}
Figure \ref{fig:mmap_mem_usage} shows the growth in memory usage of the system
for the workloads described in section \ref{ss_vm_area_struct}.

@@

The memory overhead of \texttt{vm\_area\_struct}s increases linearly with the
number of mapped regions when these regions are separated by unused memory
addresses. When a workload maps a large number of regions, this overhead can be
on the order of hundreds of megabytes to gigabytes for large memory systems.  To
reduce this overhead, it is crucial that regions with similar access permissions
are placed adjacently in the address space so that they can be identified by a
single structure. It is important for large applications such as object-oriented
databases that map a large number of regions to be aware of these overheads and mitigate them \cite{utlk}.

@@

>>>>

For demand paging to work efficiently, the correct \texttt{vm\_area\_struct}
needs to be found quickly. Linux caches the most recently used region and
organizes all regions into a red-black tree for fast lookup.  Moreover,
adjacent regions with similar permissions and properties are merged in an
effort to keep bookkeeping costs low.

%In our first set of experiments, we measure the memory overhead and latency of
%the Linux Memory Management system's use of \texttt{vm\_area\_struct}, which
%represents a region of a process's virtual address space. All the physical
%pages in the system are managed using the \texttt{struct page} descriptors in
%the kernel.  We do not measure the overhead of \texttt{struct page} because it
%is known that there is one struct for each physical page. We do not measure the
%overhead of page tables because this can be calculated from the size of the
%virtual memory space used by the process.

@@

The specifications of the machine we used to run our experiments are listed in
Figure \ref{fig:specs}. We disabled Intel SpeedStep and TurboBoost and set the scaling governor to ensure
that the processor frequency stays constant. All experiments are pinned to the
same core on our machine.

We run several different microbenchmarks, each of which has a different memory
allocation pattern. These benchmarks are designed to stress the memory
management system in different ways, rather than simulate real workloads. 
Each run is executed in isolation, without any additional processes
running beyond services that start when the system is booted and a single ssh
session. We are careful to avoid running any additional processes, which could 
impact memory usage and, thus, our results.

@@

Figure \ref{fig:mmap_mem_usage} shows the growth in memory usage of the system
for the workloads described in section \ref{ss_vm_area_struct}.

@@

The memory overhead of \texttt{vm\_area\_struct}s increases linearly with the
number of mapped regions when these regions are separated by unused memory
addresses. In the worst case, where each memory region is only one page long, \texttt{vm\_area\_struct} would represent a 5\% memory overhead per page. When a workload maps a large number of regions, this overhead can be
on the order of hundreds of megabytes to gigabytes for large memory systems.  To
reduce this overhead, it is crucial that regions with similar access permissions
are placed adjacently in the address space so that they can be identified by a
single structure. It is important for large applications such as object-oriented
databases that map a large number of regions to be aware of these overheads and mitigate them \cite{utlk}.

@@

\begin{figure}[t]
    \centering
    \begin{tabular}{c||c|c}
        & \% of ops & Time spent \\ \hline \hline
        Latency $\le$ 50ms & 99.97\% & 7.39s \\
        Latency $>$ 50ms & 0.03\% & 18.9s \\
    \end{tabular}
\end{figure}
